Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach

### 2019-08-03 Source/Thesis/Chapter1.tex
updated Paper3\

There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}. However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}. Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression. Another area of focus for research of this nature is the process of collecting appropriate language samples. Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task. Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD. There is some evidence to support this. Berisha et al \cite{Berisha2015}, has shown through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control. 

@@

Once we have a way to collect language samples, the question to consider is how we may best analyse this data. There are a number of factors that one must consider when answering the question. We look at the effectiveness of automatic speech recognition at translating speech into language, as well as the minimum quality of data that needs to be passed to any pipepline we develop. The Talk2Me app collects and then generates approximately 2000 lexico-syntactic, acoustic and semantic features (\cite Kommeili2019} as part of their data collection framework. Chapter 4 discusses the challenges connected to constructing an effective data pipeline and contains two experiments.

>>>>

There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}. However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}. Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression. Another area of focus for research of this nature is the process of collecting appropriate language samples. Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task. Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD. As is the case with other quantitative measures of cognitive ability, contrasting individuals with MCI and Early AD is challenging due to the variation in an individual's baseline speech capacity. It is therefore prudent to measure cognitive decline over time as suggested by Albert et al \cite{Albert2011}. There is some evidence to support this approach, Berisha et al \cite{Berisha2015} demonstrated through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control. 

@@

Once we have a way to collect language samples, the question to consider is how we may best analyse this data. There are a number of factors that one must consider when answering the question. We look at the effectiveness of automatic speech recognition at translating speech into language, as well as the minimum quality of data that needs to be passed to any pipepline we develop. The Talk2Me app collects and then generates approximately 2000 lexico-syntactic, acoustic and semantic features {\cite Komeili2019} as part of their data collection framework. Chapter 4 discusses the challenges connected to constructing an effective data pipeline and contains two experiments.
